# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data

```{r echo = TRUE}
library(ggplot2)
# read data from "activity.csv"
activityData <- read.csv("activity.csv", stringsAsFactors = FALSE)

# clean NA values in steps of data
cleanData <- activityData[!is.na(activityData[,1]),] 

# split data by date
splitData <- split(cleanData[,1], cleanData$date)

# sum the steps
sumData <- lapply(splitData, sum)

# tranform to frame
ssData <- data.frame(sumData)

# get the sum of steps
rData <- 0
for (i in 1 : length(ssData)){
    rData[i]<- ssData[i][,1][1]
}

# tranform to frame
stepsData <- data.frame(rData)
```

## What is mean total number of steps taken per day?
1. Make a histogram of the total number of steps taken each day
```{r echo = TRUE}
# set names of stepsDate
names(stepsData) <-"steps"

# make a histogram of the total number of steps taken each day
qplot( steps,data = stepsData)
```

2. Calculate and report the mean and median total number of steps taken per day
```{r echo = TRUE}
# summary the steps
summaryData <- lapply(splitData, summary)

# print summary
print(summaryData)
```

## What is the average daily activity pattern?
1. Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r echo = TRUE}
# average steps per day
avgData <- lapply(splitData, ave)

# get names 
namesData <- names(avgData)

# get average sum of steps
rData <- 0
nameS <- ""
for (i in 1 : length(namesData)){
    unaV <- unlist(avgData[namesData[i]])
    rData[i]<- unaV[1]
    }

# tranform to frame
stepsData <- data.frame( rData)#data.frame(rData)
stepsData["dates"]<-namesData
# set names of stepsDate
names(stepsData) <-c("steps", "dates")

# make a plot of the average number of steps taken each day
stepsData <- stepsData[!is.na(stepsData[1]),]
plot(1:length(stepsData$dates),stepsData$steps,type = "l")

```

2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r echo = TRUE}
print(max(rData[!is.na(rData)]))
print(namesData[!is.na(rData) & rData == max(rData[!is.na(rData)])])
```
## Imputing missing values
1. Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{r echo = TRUE}
naRowsNum = sum(is.na(activityData[,1]))

countAllPartNaDay <-0
# check if all intervals per day are NA
for (i in 1 : length(namesData)){  
    # get number of intervals which value is NA per day
    naValueNum<- sum(is.na(activityData[,1]) & activityData[,2]==namesData[i])
    # get number of intervals which value is not NA per day
    notNaValueNum<- sum(!is.na(activityData[,1]) & activityData[,2]==namesData[i])
    
    # List all intervals per day are NA")
    if (naValueNum > 0  & notNaValueNum >0){    
        print(namesData[i])
        print(naValueNum)
        print(notNaValueNum)
        countAllPartNaDay <- countAllPartNaDay + 1
    }
    
  }

```
The total number of missing values in the dataset is `r naRowsNum`. But check each day intervals that intervals is all NA of the day. The number of days, their part of invtervals are NA s and part are not NA s,
is `r countAllPartNaDay`. It means if intervals of one day have some NA s, the all intervals are NA s.

2. Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.
```{r echo = TRUE}
# get mean of all steps 
stepsMean <- mean(stepsData[,1])
```

3. Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r echo = TRUE}
activityDataNew <- activityData
# set stepsMean to steps of intervals which steps is NA 
activityDataNew$steps[is.na(activityDataNew$steps)] <- stepsMean
```

4. Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?
```{r echo = TRUE}

# split data by date
splitData <- split(activityDataNew[,1], activityDataNew$date)

# sum the steps
sumData <- lapply(splitData, sum)

# tranform to frame
ssData <- data.frame(sumData)

# get the sum of steps
rData <- 0
for (i in 1 : length(ssData)){
    rData[i]<- ssData[i][,1][1]
}

# tranform to frame
stepsDataN <- data.frame(rData)

# set names of stepsDate
names(stepsDataN) <-"steps"

# make a histogram of the total number of steps taken each day
qplot( steps,data = stepsDataN)

# summary the steps
summaryData <- lapply(splitData, summary)

# print summary
print(summaryData)
```

Because we use mean steps for missing data, so we can see the count of mean steps increase as the histogram shown. I think it's a Normal Distribution.

## Are there differences in activity patterns between weekdays and weekends?
1. Create a new factor variable in the dataset with two levels -- "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.
```{r echo = TRUE}
# initial
weekdaysList <- ""

# get weekdays
Sys.setlocale("LC_TIME","en_US")
weekdaysData<- weekdays(as.Date(stepsData$dates, "%Y-%m-%d"))

# get weekend or weekday
for (i in seq_along(weekdaysData)){
    w <- weekdaysData[i]
    if (w == "Sunday" | w == "Saterday")
    {
        weekdaysList[i] <- "weekend"  
    }else{
        weekdaysList[i] <- "weekday" 
    }

}
# Create a new factor variable
weekdaysFactor <- factor(weekdaysList) 
# Add to dataset
stepsData$weekdays <-weekdaysFactor
```
2. Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). 
```{r echo = TRUE}
library(lattice)
stepsData2 <- transform( stepsData, dateN = 1: length(dates))
xyplot(steps ~ dateN |weekdays , data = stepsData2, layout =c(1,2), type = "l")
```

